Visualizing probability distributions across bivariate cyclic temporal granularities

### 2021-05-01 main.Rmd
periodical to periodic

\begin{defn}\label{def:periodical}
A granularity $G$ is \textbf{periodical} with respect to a finite granularity $H$ if:
(1) $G \trianglelefteq H$; and
(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{defn}

If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$. Condition (2) in Definition \ref{def:periodical} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$. In this pattern, each granule of $H$ is shifted by $P$ granules of $G$. $P$ is called the \textbf{Period} (@Bettini2000-qk).

For example, day is periodical with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodical with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).

Granularities can also be periodical with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy]. In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.

@@

Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodical}. Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
Examples include public holidays which repeat every year, but there is no reasonably small period within which their behavior remains constant. A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf]. In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend. Examples of aperiodic granularity may also include school holidays or a scheduled event. All of these are recurring events, but with non-periodic patterns. Consequently, $P_i$ (as given in \autoref{tab:quasi}) are essentially infinite for aperiodic granularities.

>>>>

\begin{defn}\label{def:periodic}
A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
(1) $G \trianglelefteq H$; and
(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{defn}

If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$. Condition (2) in Definition \ref{def:periodic} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$. In this pattern, each granule of $H$ is shifted by $P$ granules of $G$. $P$ is called the \textbf{Period} (@Bettini2000-qk).

For example, day is periodic with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodic with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).

Granularities can also be periodic with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy]. In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.

@@

Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodic}. Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
Examples include public holidays which repeat every year, but there is no reasonably small period within which their behavior remains constant. A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf]. In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend. Examples of aperiodic granularity may also include school holidays or a scheduled event. All of these are recurring events, but with non-periodic patterns. Consequently, $P_i$ (as given in \autoref{tab:quasi}) are essentially infinite for aperiodic granularities.
